---
title: "Geo-spatial tutorial: create a simple feature"
author: "JR Ferrer-Paris"
institute: "Centre for Ecosystem Science<br/>UNSW Data Science Hub"
date: "March 2021 (updated: `r Sys.Date()`)"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to create spatial features with R package `sf` and visualise with `mapview`.
---

```{r setup, include=FALSE}
library(learnr)
library(sf)
require(leafpop)
library(leaflet)
library(mapview)

tutorial_options(
  exercise.timelimit = 60,
  # A simple checker function that just returns the message in the check chunk
  exercise.checker = function(check_code, ...) {
    list(
      message = eval(parse(text = check_code)),
      correct = logical(0),
      type = "info",
      location = "append"
    )
  }
)
knitr::opts_chunk$set(error = TRUE)


```

## Welcome

This is part 1 of a series of introductory tutorials to work with geo-spatial data in R. I prepared these tutorials as a intuitive "hands on" introduction, but I provide links for those interested in more background and theory.

In this tutorial, you will learn how to create a spatial object with package `sf` (simple features).


### Setup

I've preloaded the packages for this tutorial with

```{r eval = FALSE}
library(sf)
library(leaflet)
library(mapview)
library(leafpop)
```


For this tutorial we will **create our own spatial objects from scratch**, so we don't need to load any additional data.


```{r, eval=TRUE, child="01-create-sf-section.Rmd"}

```

### Good job!

We just created our first spatial object, how cool is that?

So we learned to:

- use package `sf`
- define a geometry with `st_points`
- assign a coordinate reference system with `st_crs`
- create a spatial object with `st_sf` by combining a data frame with a simple feature column with `st_sfc`
- combine `sf` objects with `rbind`

But what can we do with these sets of points? Let's do some more spatial stuff like calculating distance and mapping in the next section!


```{r, eval=TRUE, child="02-map-points-section.Rmd"}

```

### That's it for this tutorial!

Well, that was easy! wasn't it?

We learned to:

- calculate distance between points with `st_distance` and `st_nearest_feature`
- map two sets of points using `mapview`

I hope this has been useful and you feel more confident to start working with geo-spatial data in R.

Check the next tutorials in our [geospatial-data repository](https://github.com/UNSW-codeRs/geospatial-data-in-R)
