
## Create a grid for spatial prediction 

We read the JBM boundary file from our data folder:

```{r}
boundary_jbm <- read_sf(here::here(data_folder, "JBM-boundary-no-crs.gpkg"))
```

First we create an empty raster using functions from package `raster`.

```{r, context="setup"}
rnull <- raster(extent(boundary_jbm),
                nrows=18, ncols=24,
                crs=st_crs(boundary_jbm)$proj4string)
```

###

Then we transform this to a mask by assigning a value of 1 for cells within 50 meters of a sampling point. We use the function `distanceFromPoints`.

```{r, context="setup"}
mask <- distanceFromPoints(rnull,data_jbm)<50
```

###

Finally, we create a Spatial Pixels Data Frame for the spatial prediction. The package `gstat` uses this format inherited from the old `sp` package. 

```{r, context="setup"}
ss <- values(mask)==1
prd.grd <- SpatialPixelsDataFrame(
  points=xyFromCell(mask,1:ncell(mask))[ss,],
  data=data.frame(values(mask)[ss]),
  proj4string=CRS(st_crs(boundary_jbm)$proj4string))
```

### Visualise the mask 

Run this chunk to visualise the raster mask before continuing to the next step:
```{r mask, exercise = TRUE}
plot(prd.grd, main = "Mask") 
plot(st_geometry(data_jbm), add=TRUE)
```
