
## Kriging

In Geostatistics the term **Kriging** is used for a family of models that minimise the spatial error and can be used for spatial interpolation and prediction. Here we use **Indicator kriging** to interpolate the probability of $Z = \mathrm{Tree\ cover} > 2$ from out measurement points into a regular grid.

Here all of our previous steps come together:

```{r kriging_fit, context = "setup"}
ik = krige(Z ~ 1, # our target variable
           as_Spatial(data_jbm), # our spatial data
           prd.grd, # the regular grid for prediction 
           variogram_model) # the spatial autocorrelation

```

### Visualise the prediction

```{r kriging_pred, exercise = TRUE}
spplot(ik["var1.pred"],
       main = "indicator kriging predictions",
       sub = "Z = Tree Cover > 2")
```

### Visualise the variance


```{r kriging_var, exercise = TRUE}
spplot(ik["var1.var"],
       main = "indicator kriging variance",
       sub = "Z = Tree Cover > 2")
```

What kind of geospatial analysis is this?

What kind of geospatial data was produced in the prediction?
