---
title: "Geo-spatial tutorial: spatial interpolation"
author: "JR Ferrer-Paris"
affiliation: "Centre for Ecosystem Science / UNSW Data Science Hub"
date: "March 2023 (updated: `r Sys.Date()`)"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to fit a variogram and do spatial interpolation using R package `gstat`.
bibliography: references.bib
---

```{r setup, include=FALSE}
library(learnr)
library(sf)
library(gstat)
library(dplyr)
library(raster)
library(ggplot2)

tutorial_options(
  exercise.timelimit = 60,
  # A simple checker function that just returns the message in the check chunk
  exercise.checker = function(check_code, ...) {
    list(
      message = eval(parse(text = check_code)),
      correct = logical(0),
      type = "info",
      location = "append"
    )
  }
)
knitr::opts_chunk$set(error = TRUE)
here::i_am("workshop/tutorial-4/tutorial-4-geostatistic-example.Rmd")
data_folder <- "data"

# prepare data
newdata <- here::here(data_folder,"JBM-points-no-crs.gpkg")
olddata <- here::here(data_folder,"JBM-points.gpkg")
if (!file.exists(newdata)) {
    dts <- read_sf(olddata)
    data_jbm <- dts #%>% mutate(TCC=factor(`Tree cover`))
    st_geometry(data_jbm) <- st_geometry(data_jbm) - c(203000,1171000)
    st_crs(data_jbm) = st_crs('LOCAL_CS["Undefined Cartesian SRS"]')
    write_sf(data_jbm, dsn=newdata)
}
data_jbm <- read_sf(newdata)
st_crs(data_jbm) = NA

newdata <- here::here(data_folder,"JBM-boundary-no-crs.gpkg")
olddata <- here::here(data_folder,"JBM.gpkg")
if (!file.exists(newdata)) {
  dts <- read_sf(here::here(data_folder, "JBM.gpkg"))
  boundary_jbm <- filter(dts, id %in% c(7,8,11))
  st_geometry(boundary_jbm) <- st_geometry(boundary_jbm) - c(203000,1171000)
    st_crs(boundary_jbm) = st_crs('LOCAL_CS["Undefined Cartesian SRS"]')
    write_sf(boundary_jbm, dsn = newdata)
}
boundary_jbm <- read_sf(newdata)
st_crs(boundary_jbm) = NA

xys <- st_coordinates(data_jbm)
distance_matrix <- st_distance(data_jbm)

```

## Welcome

This is part of a series of introductory tutorials to work with geo-spatial data in R. I prepared these tutorials as a intuitive "hands on" introduction, but I provide links for those interested in more background and theory.

In this tutorial, you will follow some steps to read spatial data in vector format with package `sf` (simple features), estimate the spatial correlation of variables associated with these points and use a semivariogram model to apply spatial interpolation following the indicator kriging approach.

### Setup

I've preloaded the packages for this tutorial with

```{r eval = FALSE}
library(sf)
library(gstat)
library(ggplot2)
library(raster)
```


For this tutorial we will **read spatial objects from a file**. This file is preloaded in this `learnr` tutorial, but if you want to replicate this steps in an R session, you need to locate the original file in the data folder of the parent repository. For example, using package `here`:

```{r eval = FALSE}
here::i_am("workshop/tutorial-4/tutorial-4-geostatistic-example.Rmd")
data_folder <- here::here("data")
# Check if file exists at data folder
file.exists(here::here(data_folder, "JBM-points.csv"))
file.exists(here::here(data_folder, "JBM.gpkg"))
```

```{r, eval=TRUE, child="01-data-exploration.Rmd"}

```

### One step done...

So far we have

- read our data using package `sf`

We are just getting started, let's keep going!

```{r, eval=TRUE, child="02-visualise-data.Rmd"}

```

### This looks good!

So far we have

- read our data using package `sf`
- used `ggplot2` for visualising one variable

But we want to quantify the spatial autocorrelation of the data and try to make a spatial prediction for this variable. Let's figure out how to do this in the next steps!

```{r, eval=TRUE, child="03-spatial-autocorrelation.Rmd"}

```

### Just mastered spatial auto-correlation!

So far we have

- read our data using package `sf`
- used `ggplot2` for visualising one variable
- used `gstat` functions for fitting a variogram model

Now we are almost ready to make a spatial prediction, but first let's create a raster mask!

```{r, eval=TRUE, child="04-create-grid.Rmd"}

```

### From vector to raster!

Awesome! So far we have:

- read our data using package `sf`
- used `ggplot2` for visualising one variable
- used `gstat` functions for fitting a variogram model
- used `raster` and legacy `sp` functions to create a spatial grid for prediction

Here comes the fun part: spatial prediction with kriging.

```{r, eval=TRUE, child="05-kriging.Rmd"}

```

## That's it for this tutorial!

Good job!

We learned to:

- read our data using package `sf`
- used `ggplot2` for visualising one variable
- used `gstat` functions for fitting a variogram model
- used `raster` and legacy `sp` functions to create a spatial grid for prediction
- used `gstat` functions for spatial prediction (kriging)

I hope this has been useful and you feel more confident to start working with geo-spatial data in R.

### What is next?

Check other tutorials in this workshop at the [UNSW codeRs' geospatial-data repository](https://github.com/UNSW-codeRs/geospatial-data-in-R) or in the following links:

```{r echo=FALSE, results='asis'}
cat(knitr::knit_child(text = readLines(here::here("tutorials-links.Rmd")), quiet=TRUE), sep = '\n')
```

### References

The data for this tutorial comes from:

<div id="refs"></div>
