
## A first thematic map

A thematic map is a map that displays or visualises some aspects of the data, not just the spatial boundaries. We have seen that the World dataset includes multiple columns with information for each country.

Run this code to create your first thematic map:

```{r first_thematic_map, exercise=TRUE}
tm_shape(World,projection='+proj=robin') + 
  tm_polygons(
    col = "income_grp", 
    palette = "-Blues", 
    contrast = 0.7, 
    border.col = "gray30")
```

You can replace "income_grp" with other categorical variables in your dataset and click the run button again to create a different view of your dataset. 

### What about quantitative variables ?

Functions in `tmap` will usually find a way to display quantitative data, but sometimes default parameters will not be sufficient for data with large ranges or skewed distributions. 

For example, the population estimates do not look good with the defaults. Try defining breaks, or changing the style parameter (or both), and see how the plot improves.

```{r second_thematic_map, exercise=TRUE}
tm_shape(World,projection='+proj=robin') + 
  tm_polygons(
    col = "pop_est", 
    palette = "-Blues", 
    #style = "log10_pretty",
    #breaks = c(0,1^6,10^7,10^8,10^10),
    contrast = 0.7, 
    border.col = "gray30")
```

### Title for the legend

You can use the parameter `title` to customise your legend:
```{r}
tm_shape(World,projection='+proj=robin') + 
  tm_polygons("income_grp", palette = "-Blues", 
              title = "Income class", contrast = 0.7, border.col = "gray30",
              id = "name") 
```

