---
title: "Geospatial tutorial - section 5"
author: "JR Ferrer-Paris"
---
## Interactive maps with `mapview`

Mapview is ...

Most of the examples here are based on the help-page and tutorial of mapview, I encourage you to explore their examples to learn more about the full range of options.

###
Mapview uses a similar syntax as `ggplot`, you can add components with the `+` operator:

```{r}
mapview(franconia) + mapview(trails) + mapview(breweries) + mapview(UBT)
```

Navigation of a mapview element is quite intuitive

- Controls allow you to zoom and pan,
- Use the layer icon to switch layers on and off
- You can also change the background layers
- Click on an element to read its table of attributes
- There are shortcuts to zoom to each layer or for a full zoom

###

We can change the colors and other attributes of each layer


```{r}
UBT_url <- "https://www.study-in-bavaria.de/fileadmin/_processed_/csm_Campus-Luftbild_1_qf_03_8a455a1f25.jpg"

mapview(franconia, zcol = "district", legend = TRUE) + 
  mapview(trails) + 
  mapview(breweries, zcol = "founded", at = seq(1400, 2200, 200), legend = TRUE) + 
  mapview(UBT, popup = popupImage(img=UBT_url, src ="remote"), col.regions = "blanchedalmond") 
```

###

You can also combine several function calls to create maps of computed values for your spatial objects. This is an example for creating a map of density of breweries per municipality

```{r}
franconia %>%
    mutate(count = lengths(st_contains(., breweries)),
           area = st_area(.) %>% set_units('km^2'),
           density = count / area) %>%
    mapview(zcol = "density")

```

