---
title: "Geo-spatial tutorial: thematic map"
author: "JR Ferrer-Paris"
date: "March 2021 (updated: `r Sys.Date()`)"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to ...
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(tmap)
library(sf)

tutorial_options(
  exercise.timelimit = 60,
  # A simple checker function that just returns the message in the check chunk
  exercise.checker = function(check_code, ...) {
    list(
      message = eval(parse(text = check_code)),
      correct = logical(0),
      type = "info",
      location = "append"
    )
  }
)
knitr::opts_chunk$set(error = TRUE)
data("World", "metro", package = "tmap")
st_crs(metro) = "EPSG:4326"
metro <- 
  metro %>% 
  mutate(growth = (pop2020 - pop2010)/(pop2010 * 10) * 100)


```

## Welcome

In this tutorial, you will learn some basic steps to start working with (geo)-spatial data in R. I prepared this tutorial as a intuitive "hands on" introduction, but I provide links for those interested in more background and theory.

We will cover following steps :

* Creating a thematic map with vector data (shapes)
* Combining multiple layers from one or multiple shapes
* Switching between plot and interactive maps

### Setup

I've preloaded the packages for this tutorial with

```{r eval = FALSE}
library(tmap)
library(sf)
```

Throughout this tutorial we will be using the  pipe operator, `%>%`. You can use the pipe to rewrite multiple operations in a way that you can read left-to-right, top-to-bottom. We'll use piping frequently because it considerably improves the readability of code. The pipe is a defining feature of the tidyverse, so we will load this set of packages as well.

```{r eval = FALSE}
library(tidyverse)
```

The tidyverse package also includes `ggplot2`, which allow us to create static maps in a similar modular fashion, but using the `+` operator instead of the pipe.

For this tutorial we will use the `World`, `metro` and other data sets from package `tmap`. These data sets are documented in `help(package=tmap)`. 

Examples in this tutorial are based on:

> Tennekes, M. (2018). tmap: Thematic Maps in R. Journal of Statistical Software, 84(6), 1–39. [DOI:10.18637/jss.v084.i06](https://doi.org/10.18637/jss.v084.i06)


```{r, eval=TRUE, child="00-read-data.Rmd"}

```

###

Now that we have a basic understanding of the data at hand, let's use this information to create thematic maps.


```{r, eval=TRUE, child="01-thematic-map.Rmd"}

```

###

Great! we created a thematic map displaying information from one variable in our dataset. Often we want to combine multiple layers to be able to see connections between our variables. In the next section we will see how to make this work with `tmap`.


```{r, eval=TRUE, child="02-multiple-layers.Rmd"}

```

###

This looks nice! Our map is loaded with information, but the world is a big place, and the details are sometimes hidden in tiny places. How can we dive into the map and explore every corner of it? 

In the next section we will turn this static plot into an awesome interactive map!


```{r, eval=TRUE, child="03-interactive-map.Rmd"}

```

### That's it from me!

We just learned to create thematic maps with vector data in `tmap`, combining multiple layers from one or multiple shapes and switching between static plot and interactive maps.

`tmap` includes many layers, functions and features that can help you create beautiful maps for your presentations and publications. You can get inspired by this final example:

```{r}

data("land", "rivers", package = "tmap")
m2 <- tm_shape(land) +
  tm_raster("elevation", breaks = c(-Inf, 250, 500, 1000, 1500, 2000, 2500, 3000, 4000, Inf),  
            palette = terrain.colors(9), title = "Elevation (m)") +
  tm_shape(rivers) + 
  tm_lines("lightblue", lwd = "strokelwd", scale = 1.5, legend.lwd.show = FALSE) +
  tm_shape(World, is.master = TRUE,projection = "+proj=eck4") +
  tm_borders("grey20", lwd = .5) +
  tm_grid(projection = "+proj=longlat", labels.size = 0.4, lwd = 0.25) +
  tm_text("name", size = "AREA") +
  tm_compass(position = c(0.08, 0.45), color.light = "grey90", size = 3) +
  tm_credits("Eckert IV projection", position = c("RIGHT", "BOTTOM")) +
  tm_style("classic",
           bg.color = "lightblue",
           space.color = "grey90",
           inner.margins = c(0.04, 0.04, 0.03, 0.02), 
           earth.boundary = TRUE) +
  tm_legend(position = c("left", "bottom"), 
            frame = TRUE,
            bg.color = "lightblue")
tmap_mode("plot")
m2

```

I hope this has been useful and you feel more confident to start working with geo-spatial data in R. 

Keep exploring the world!
