---
title: "Geospatial tutorial - section 2"
author: "JR Ferrer-Paris"
---
## Explore points with `sf`

For this tutorial we want to explore the surroundings of the University of Bayreuth, my _alma mater_.

I created a simple feature object using the functions we learned in the previous tutorial:

```{r UBT}
UBT <- st_sf(data.frame(full_name='Universität Bayreuth',
             url='https://uni-bayreuth.de/'),
             crs=st_crs("EPSG:4326"),
             geometry=st_sfc(
               st_point(c(11.585833, 49.928889))))

```

### Welcome to Franconia

There is not much we can do with one point, so we will look at a couple of data sets from this region, known as "Franken" in german or "Franconia" in other languages.

### The _breweries_ dataset in Franconia

The best way to start our tour is to explore one of its main attractions: the breweries or Brauereien.

If you want to have an overview of a `sf` object in R, just type its name, click on the run button to show data and attributes of this object:

```{r pointsex1, exercise=TRUE, paged.print=FALSE}
breweries
```

This is basically the same structure we have seen before, but with more points and several variables.

If you get a warning about and old-style crs object, you can simply update this with:

```{r}
st_crs(breweries) <- "EPSG:4326"
```

### Separate data from geometry

We can use the function `st_drop_geometry` to look at the data without the spatial component:

```{r pointsex2, exercise=TRUE}
breweries %>% st_drop_geometry()
```

Or we can use `st_geometry` to look at the spatial component without the data:
```{r pointsex3, exercise=TRUE, paged.print=FALSE}
breweries %>% st_geometry()
```

### Plotting geometry and data

There are many ways to explore data in spatial objects. We might be interested in non-spatial information, for example, how many types of beers are there in each brewery?

```{r pointsex4, exercise=TRUE}
breweries %>% pull(number.of.types) %>% hist(main="Nr. of types of beers")
```

###
But how do we know where are the breweries with the most types of beers?

Here we will select one variable from our object and use the default plot function for _sf_ objects:

```{r pointsex5, exercise=TRUE}
breweries %>% select(number.of.types) %>% plot(pch=19)
```

So, where would you go to have the largest selection of beers?

### Distance between spatial objects
You can imagine this is a question that was frequently asked in academic circles of the University. So let's find the answer!

- First we will need to define what is the desired number of beers we expect. We can do this by filtering our spatial object based on a variable:

  ```{r eval=FALSE}
  selected_breweries <- breweries %>% filter(number.of.types>9)
  ```

- Then we want to know which one of these is nearest to our point of reference (UBT point):

  ```{r eval=FALSE}
  qry <- st_nearest_feature(UBT,selected_breweries)
  ```

- Now we can check the details for this record in the spatial object:

  ```{r eval=FALSE}
  selected_breweries %>% slice(qry)
  ```

- Finally we want to know the distance from the uni:

  ```{r  eval=FALSE}
  st_distance(UBT,selected_breweries %>% slice(qry))
  ```

###

Try running these lines of code, and see if you can answer the questions below:

```{r  pointsex6, exercise=TRUE}
selected_breweries <- breweries %>% filter(number.of.types>9)
qry <- st_nearest_feature(UBT,selected_breweries)
selected_breweries %>% slice(qry)
st_distance(UBT,selected_breweries %>% slice(qry))
```

```{r how_many_beers, echo = FALSE}
question(" What's the name of the place, How many types of beers will we find there? and how far do we have to travel? ",
         answer("Brauerei Huetten, 10 km and 14808 beers", message = "Ok, you have had too many beers, please try again"),
         answer("Kitzmann-Braeu, 14 beers and 55 km", message = "Did you change the threshold? this is a far ride, but gets you the largest variety of beers"),
          answer("Brauerei Huetten, 10 beers and 14.8 km", correct = TRUE),
         answer("Brauhaus Leikeim, 11 beers, 33km ", message = "Did you change the threshold? A good choice if you really need more than ten beers."),
         allow_retry = TRUE
)
```
