---
title: "Geo-spatial tutorial: create a simple feature"
subtitle: "Section 3"
author: "JR Ferrer-Paris"
---

In the last section I will give some advice on how to import data from different file formats.


```{r, eval=FALSE, child="03-import-section.Rmd"}

```

## Importing spatial data into R

There are plenty of options for importing spatial data into R. Here I show only a few options, please follow the links to read more advanced documentation:

### Load existing datasets

If you need basic layers of you can load data from existing R packages, for example:

***rnaturalearth***: Vector map data from [natural earth](http://www.naturalearthdata.com/)

```{r}
library("rnaturalearth")
library("rnaturalearthdata")
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)
head(world,1)
```

***ozmaps***: Maps of Australian coastline and administrative regions. Data can be drawn or accessed directly as simple features objects.

```{r}
library(ozmaps)
sf_oz <- ozmap_data("states")
class(sf_oz)
head(sf_oz,1)
```

### Importing data using `st_read`

The function `st_read` allows to read spatial data from many different formats. For example, you can read a geojson from a text string:

```{r, eval=FALSE}
geojson_txt <- paste("{\"type\":\"MultiPoint\",\"coordinates\":",
   "[[3.2,4],[3,4.6],[3.8,4.4],[3.5,3.8],[3.4,3.6],[3.9,4.5]]}")
x = read_sf(geojson_txt)
x
```

To read from a shapefile, you need to specify the path to the one its components (either file ".shp", ".shx" or ".dbf" )

```{r, eval=FALSE}
my_shape = st_read("path-to-your-file/filename.shp")
summary(my_shape)
```

To read from a postGIS or similar spatial database, you need to set up a connection to the database and then select the table or query you want to execute in the database:

```{r, eval=FALSE}
library(RPostgreSQL)
conn <- dbConnect(PostgreSQL(), dbname = "postgis")
x = st_read(conn, table = "public.meuse")
dbDisconnect(conn)


```

We just learned to:

- load data from existing packages,
- read geojson and shapefiles with `st_read`
- read from postGIS using `st_read` and a database connection
